---
title: "Meeting Notes"
author: "Joshua delabruere"
date: "2/15/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

First we begin by calling libraries that we want to use. 
I'm loading the whole tidyverse which includes several libaries like 

* ggplot2: Graphics
* dplyr: data manipulation (SQL like)
* tidyr: data tidying/ organizing
* stringr: for string operations
```{r cars}
library(tidyverse)
library(tidymodels)
library(randomForest)
library(factoextra)
```


Next, we should load the data. Today we're looking at the wine dataset which we found at https://www.kaggle.com/brynja/wineuci 

I've loaded the data on the shared drive


I've opted to use the readr packages ```read_csv()``` function. It has faster read time and leaves more things as characters which avoids changing variables like dates before we decide on a format. This function takes a filepath and reads the data, returning a *data frame* which is matrix like in structure. 
```{r}
wine <- readr::read_csv("wine.csv", col_names = F)
head(wine)
```


This data set came without headers, so we'll have to add column names by themselves. We can do this by creating a *vector*. A vector is a list, but all items must be the same type. A data frame is composed of vectors of the same length. Each vector represents a variable in a data frame. 

we can create a vector using ``` c(item1, item2, ... itemn) ``` 

Remember in R we use ``` <- ``` instead of `=` for variable asignment like ```variable <- 1+4```. We can use ```=``` for assignment, but R best practices discourages this and reserves ```=``` for equivelence checks. 

variables names are as follows:

* cultivar: The varaity of wine
* alcohol: 
* malic_acid: 
* ash 
* alcalinity_ash
* magnesium 
* total_phenols 
* flavanoids 
* non_flavnoid_phenols
* proanthocyanins
* color_intensity
* hue
* od
* proline

We can also use the function ```names(data_frame_name)``` to access the column names of a data frame. We will assign our list of variable names to the names of our data frame. 

```{r}
varNames <- c('cultivar', 'alcohol', 'malic_acid', 'ash', 'alcalinity_ash', 'magnesium', 'total_phenols', 'flavanoids', 'non_flavnoid_phenols', 'proanthocyanins', 'color_intensity', 'hue', 'od', 'proline')

names(wine) <- varNames

head(wine)

```

Next lets get the summary statistics of our data set.  We can use the ```summary()``` command to get summary statistics.

```{r}
summary(wine)
```
Notice that cultivar is being treated as a number rather than a factor. We can fix that using ```as.factor()```. We can directly access a column in the data frame using ```df$varName``` so for us, ```wine$cultivar``` **Or** we can use ```wine[,1]```


```{r}
wine$cultivar <- as.factor(wine$cultivar)
summary(wine)
```

We can use data visualization to examine the relationships between various variables. I used a heatmap to look at how correlated the various columns were. for correlation we need all columns to be numeric, so let's exclude column 1 from our data set using ```wine[,-c(1)]``` we can remove the values I've set to na for slight clustering with dendograms

```{r}
heatmap(cor(wine[,-c(1)]), Rowv = NA, Colv = NA)
heatmap(cor(wine[,-c(1)]), Rowv = NA)
heatmap(cor(wine[,-c(1)]), Colv = NA)
heatmap(cor(wine[,-c(1)]))
```

Looking at the darker squares we can see which variables are closely related. 


We can also plot the data set to create a cor plot using ```plot(wine)```

Play around with the variables you include as this is pretty cluttered. You can exclude variables by putting a negative vector ```-c(1,2,3)``` exculudes the first 3 columns. You can also specifically include variables by using a regular vector ``` c(4,5,6)``` includes the columns 4,5, and 6. 

```{r}
plot(wine)
```


We want to classify today, and the easiest column to do that with is cultivar. However, we have 3 categories, which while possible, is a little more indepth than a binary classification. lets use dplyr functions to create a new variable wine_type where we're trying to identify wines from cultivar #2

dplyr (and the rest of tidyverse) use the *`%>%`* operator which is refered to as piping to pass a *data frame* like object called a *tibble* between functions which have the first argument of data and return a data frame/tibble. I think this is easier to understand in practice, so look at the example below. For this problem we're using ```mutate()``` to create a new variable and  ```select()``` to select columns.  

For simplicity sake in the example, lets select only 3 predictors: flavanoids, hue, and alcohol


```{r}
wine %>%
  mutate(wine_type = if_else(cultivar==2, 1,0)) %>%
  select(wine_type, flavanoids, hue, alcohol) %>%
  na.omit()-> classification_df

head(classification_df)
```

let's plot some scatterplots to look at how the seperation of our groups looks. I'm using ggplot 2, where you set asthetics in aes()
```{r}
classification_df %>%
  ggplot(aes(x=flavanoids, y=alcohol, color=as.factor(wine_type)))+
  geom_point()+
  theme_minimal()+
  theme(legend.position = "bottom")

classification_df %>%
  ggplot(aes(x=flavanoids, y=hue, color=as.factor(wine_type)))+
  geom_point()+
  theme_minimal()+
  theme(legend.position = "bottom")

classification_df %>%
  ggplot(aes(x=hue, y=alcohol, color=as.factor(wine_type)))+
  geom_point()+
  theme_minimal()+
  theme(legend.position = "bottom")
```


Lets's split the data into a training set and validation set. We can use tidyverse functions for this as well. 


```{r}

summary(classification_df)
classification_df$wine_type <- as.factor(classification_df$wine_type)

classification_df %>%
  mutate(id = row_number())-> classification_df


set.seed(42)
rsample::initial_split(classification_df, prop = .6)-> wine_split
rsample::training(wine_split) -> train
rsample::testing(wine_split) -> test


# classification_df %>%
#   sample_frac(.6) -> train
# 
# classification_df %>%
#   anti_join(train, by= 'id') -> test

length(train$id)
length(test$id)


```




## Lets do logisitic regression to classify


```{r}


our_logit <- logistic_reg( mode= "classification") %>%
  fit(wine_type~.,data = train)

#ourlogit <- glm(wine_type~flavanoids+alcohol+hue, data = train, family=binomial())

summary(our_logit)

```

Lets test how well our model preforms
```{r}
predictions_glm <- our_logit %>%
  predict(new_data = test) %>%
  bind_cols(test %>% select(wine_type))



predictions_glm %>%
  conf_mat(wine_type, .pred_class) 

```


Specificity, sensitivity can be calulated from this table

sensiticity is (TP/(TP+FN))
```{r}
44/47
```

Specificity = (TN/(TN/FP))
```{r}
24/(24)
```




## Lets do Random Forest







```{r}
set.seed(42)
our_rf <- rand_forest(mode="classification", trees = 2000) %>%
  set_engine("ranger", importance = "impurity") %>%
  fit(wine_type~., train)


summary(our_rf)

prediction_RF <- predict(our_rf, new_data = test) %>%
    bind_cols(test %>% select(wine_type))

prediction_RF %>%
  conf_mat(wine_type, .pred_class)






```


#Cart

```{r}

our_cart <- decision_tree(mode="classification") %>%
  fit(wine_type~., train)

prediction_cart <- our_cart %>%
  predict(test) %>%
  bind_cols(test %>% select(wine_type))

prediction_cart %>%
    conf_mat(wine_type, .pred_class)

```
